\fontfamily{phv}\selectfont{Research A Proposal}

### 2013-09-24 docs/proposal/doc.tex
Added small part on problem statement

The field of brain connectivity finds its roots in the early 1990s \cite{friston1993functional, friston1994}, though because of more recent developments in the field of Artificial Intelligence and Machine Learning it is possible to do more thorough analysis \cite{vandenheuvel2010}.
One type of connectivity analysis involves finding causal relations between brain areas.
However, causal analysis is also relatively new and its application in brain connectivity analysis still faces many challenges that require further research \cite{Ramsey2010}.

\paragraph{Problem statement}
This project seeks to combine the strengths of current state-of-the-art causal discovery methods and apply them to resting state brain fMRI data in order to find new applicable methods for determining causal patterns in the human brain.

@@

\subsection*{Brain connectivity}
\subsubsection*{Functional connectivity}
The functional connectivity describes the statistical dependance of neuronal activity of different brain regions \cite{friston1993functional}.
This gives an insight in the organization of the brain, even though the exact functioning is still unclear \cite{Lowe2000, doria2010, Bullmore2009}.
Functional connectivity is strongly time-dependant, as these dependancies may be short-lasting.

%The Functional connectivity is defined as the temporal dependence of neuronal activity patterns of anatomically separated brain regions \cite{friston1993functional}.
%An increasing body of studies successfully investigates functional connectivity by applying statistical methods on the level of co-activation of resting-state fMRI time-series between brain regions.
%Several studies have shown that resting-state networks reflect brain connectivity in a concise way, though their exact relation remains unknown \cite{Lowe2000, doria2010, Bullmore2009}.
%Resting state indicating that subjects are instructed to relax without thinking of something in particular, to stimulate spontaneous brain activity.

\subsubsection*{Structural connectivity}
The structural connectivity is a mapping of the physical paths in the brain, and it shows the neural connections between anatomically separated brain regions.
This is strongly related with functional connectivity: regions can only be functionally connected if there is a physical connection between them.
This is an intuitive relation, and is also shown experimentally \cite{vandenheuvel2009}.

@@

%Much progress has recently been made in the fields of structural and functional connectivity\cite{vandenheuvel2010}.
%Research has also been done on inferring effective connectivity, based on Granger causality \cite{Roebroeck2005}, multivariate autoregressive modeling \cite{Harrison2003} and dynamic causal modelling \cite{Friston2003}.
%However, many challenges are still being faced \cite{Ramsey2010}. One example of this is that not all approaches have a good schaling, and as such they cannot be applied to large datasets. Another problem is that not all methods can cope with latent variables, and wrong directive relations can be concluded.

>>>>

The field of brain connectivity finds its roots in the early 1990s \cite{friston1993functional, friston1994}, though because of more recent developments in the field of Artificial Intelligence and Machine Learning it is possible to do more thorough analysis \cite{vandenheuvel2010}.
One type of connectivity analysis involves finding causal relations between brain areas, specifically \emph{effective connectivity}.
However, causal analysis is also relatively new and its application in brain connectivity analysis still faces many challenges that require further research \cite{Ramsey2010}.

\paragraph{Problem statement}
This project seeks to combine the strengths of current state-of-the-art causal discovery methods and apply them to resting state brain fMRI data in order to find new applicable methods for determining causal patterns in the human brain.
Several methods for finding effective connectivity already exist \cite{mcIntosh1994, Harrisson2003, Friston2003, Roebroeck2005}
However using more generally applicable causal discovery methods has not been elaborated on much.

We would like to know whether applying such methods, such as the approach introduced in \cite{claassen2012}, or PC-algorithm introduced in \cite{spirtes2000}, can solve some of the difficulties effective connectivity suffers from.
% What problems? See \cite{Ramsey2010}.

@@

\subsection*{Brain connectivity}
\subsubsection*{Functional connectivity}
Functional
%The Functional connectivity is defined as the temporal dependence of neuronal activity patterns of anatomically separated brain regions \cite{friston1993functional}.
%An increasing body of studies successfully investigates functional connectivity by applying statistical methods on the level of co-activation of resting-state fMRI time-series between brain regions.
%Several studies have shown that resting-state networks reflect brain connectivity in a concise way, though their exact relation remains unknown \cite{Lowe2000, doria2010, Bullmore2009}.
%Resting state indicating that subjects are instructed to relax without thinking of something in particular, to stimulate spontaneous brain activity.

\subsubsection*{Structural connectivity}
The structural connectivity is a mapping of the physical paths in the brain, and it shows the neural connections between anatomically separated brain regions.
This is strongly related with functional connectivity: regions can only be functionally connected if there is a physical connection between them. \cite{vandenheuvel2009}

@@

%Much progress has recently been made in the fields of structural and functional connectivity\cite{vandenheuvel2010}.
%Research has also been done on inferring effective connectivity, based on Granger causality \cite{Roebroeck2005}, Structural equation modelling \cite{mclntosh1994}, multivariate autoregressive modeling \cite{Harrison2003} and dynamic causal modelling \cite{Friston2003}.
%However, many challenges are still being faced \cite{Ramsey2010}. One example of this is that not all approaches have a good schaling, and as such they cannot be applied to large datasets. Another problem is that not all methods can cope with latent variables, and wrong directive relations can be concluded.
